Foundational claims of group polarization are plausibly false, with lessons for the empirical study of psychological dynamics

### 2024-04-13 GroupPolarizationStatmod/paper/ms.qmd
wrote paragraph explaining selection criteria

More about group polarization history and evidence; why these ten studies and how they represent prominent theories (see Rupert Brown book, Isenberg review; Sunstein highly prominent and elevated GP in trade books)

>>>>

In this study, we take the first step towards a new kind of meta-analysis needed to demonstrate that published detections of group polarization are plausibly false. We selected sixty foundational group polarization detections from ten studies. We chose the ten studies guided by a desire to identify foundational studies that contribute to the general consensus in social psychology that group polarization reliably occurs across experimental settings—one author of one of the ten journal articles is a law professor who has written two trade books on group polarization to argue that group polarization occurs in real-world settings in addition to the laboratory, with political and other practical importance [@Sunstein2009; @Sunstein2019]. Our criteria were that the studies were either (1) written by influential authors; (2) had an outsize influence on the study of group polarization in terms of citations and journal prestige; or (3) are well-cited studies that were designed to support one of four recognized theoretical explanations of group polarization. The studies must also have used an ordinal scale for measuring participant opinions (all but a handful of exceptions use ordinal measurements to measure group polarization). Follow-on studies could establish more objective standards for inclusion, but the above criteria have provided a useful selection of ten diverse studies as we explain in more detail now.

The first prominent group polarization studies of broad scientific and social interest examined . In tracing the historical development of group polarization, four theoretical explanations have been identified that could possibly be compatible, with some practitioners considering them incompatible. Some influential studies we identify We have suggested elsewhere that a simpler explanation could go farther [@Turner2020], but here we are not interested in evaluating or comparing the these theories, but instead selecting influential studies from different recognized theoretical traditions for our selective meta-analysis here.
